Natural Selection Across Three Generations of Americans

### 2024-03-04 response-to-reviewers-revision-1.qmd

---
title: "'Natural Selection Across Three Generations of Americans': response to reviewers"
format: docx
editor: 
  markdown: 
    wrap: 72
bibliography: bibliography-hrs.bib
---

Dear Professor Medland,

Thank you for the opportunity to revise our manuscript. Thank you also
to the reviewers. We respond to their points below.

One major change is that I have asked Tobias Edwards to be a co-author.
I discovered that he was working on similar topics, so we merged our
efforts. One effect is just that we have more polygenic scores to
analyse. The original results are not much changed, which is
encouraging. Another is that we look in more detail at two extra
generations, using children and grandparents of the respondents. We've
given the paper a title which reflects that. Lastly, we have more
careful and accurate estimates of effect sizes of "true" and/or
"chip-wide" scores.

I hope this is OK. I don't want to bait and switch. But doing this seems fair to Tobias, and I think it has also improved the paper.

@@

In the previous manuscript we used a multivariate model, controlling for
genetic principal components. We now no longer control for the PCs. Now
we residualise the polygenic scores on the PCs before standardizing the
polygenic scores. This ensures our regression estimates are exactly the
covariance between the polygenic scores and RLRS.

*The authors might want to discuss the expected selection in terms of
relevance. They quantify it but they do not discuss these
quantifications. This is very important shaping our expectation of
ongoing natural selection. It might also be very interesting to know
more about whether those trends follow the secular trends in
phenotypes.*

We have now elaborated on the error corrected results and added a final
paragraph to the Discussion section on this. We estimate how selection
against intelligence will affect the phenotype, finding the result is
not especially large. In particular we discuss PGIs for health
phenotypes, as a corrective to the existing literature which is perhaps
a bit over-focused on cognition and IQ.

We briefly mention that genetic change in traits is often opposite to
the phenotypic change. This highlights that environmental change is
likely to be of greater importance for phenotypic trends than genetic
change.

@@

*Finally, the statement 'in the UK Biobank sample, correlations of PGS
coefficients between younger and older parents are negative (-0.319),
and for 42 out of 58 PGS, coefficients have opposite signs. However, in
bootstraps, these statistics are not significantly different from 0 and
29 respectively.' was not entirely clear to me. What does it mean that
42 PGS had opposite sign effects? Does that mean for old versus younger
parents' coefficients were very different, even in the opposite
direction? How does this relate to the first part of the sentence that
coefficients were negative. Also, why has this been tested against 0 and
29?*

The earlier claim is not "coefficients are negative overall" but
"coefficients among younger parents are negatively correlated with
coefficients among older parents". We are dealing with correlations of
correlations: this can get confusing.

Negatively correlated, and more strongly, opposite-signed, effects
between young and old parents support the theory that income effects
dominate for older parents, while substitution effects dominate for
younger parents. To be a bit more explicit: if a PGS simply doesn't
correlate with human capital, we expect it to not be selected on among
either group of parents. If a PGS is correlated with human capital, that
will lead to income and substitution effects for those with more human
capital. If the substitution effect dominates for younger parents then
the PGS will have a negative correlation with fertility; if the income
effect dominates for older parents, then the PGS will have a positive
correlation with fertility.

@@

However, thinking about this issue a bit more made us change our tests
more generally. The original situation was that we tested most pairs of
coefficients across sample subgroups by comparing their absolute size.
Coefficients for older and younger parents had similar absolute sizes,
but we noticed they were opposite-signed and hence ran the tests above.
This is not very satisfactory: it's too like special pleading, and it
also suggests the original tests weren't adequate.

We now improve our tests in table 1 as follows. We take the null
hypothesis to be "selection coefficients are the same in the different
subgroups, for all PGIs". Under the null, the true difference between
selection coefficients is zero; the difference in our sample (and in
bootstrapped subsamples) should then not be significantly different from
zero. Our alternative hypothesis is that, for at least some scores,
selection coefficients against human capital are smaller among the
richer/more educated/older/married group. That is, if the selection
coefficient among the poorer and richer groups are $x_p,x_r$, and
$sign(corr)$ is the sign of the PGI's correlation with human capital,
then

$x_r \times -sign(corr) < x_p \times -sign(corr)$

Under the null, we'd expect the above equation to hold in the sample 
for just 50% of the scores. If it holds for significantly more than 50% of 
the scores (using bootstrapped C.I.s) then we can reject the null in favour 
of the alternative.

We also try a simpler approach. Using the PGIs with the
largest and most precisely estimated selection effects in the whole
sample, we estimate selection effects in different subgroups and test
for a significant difference between them via a simple interaction
term. We report these results too.

@@

Done. Indeed, the black sample is really too small to say much.

*Also, there seem to be many associations of PGS with RLRS (Figure 4),
such as ADHD, diabetes, etc., which I think are interesting findings,
even if of low significance, worth at least a brief discussion.*

We agree. We now discuss some health-related traits in the conclusion:

> Whilst past research has focused on \[cognitive\] traits, we find
> selection differentials are often of similar magnitudes for health
> related traits. The selection differential for self-rated health is
> greater than those of education and cognitive performance. The most
> significant, positively selected trait was ADHD, which was also found
> in the UK [@hugh2022human]. Future research should study the health
> and medical implications of natural selection, in addition to its
> social implications.

@@

*What is the data base for the mean number of children of people born in
the same year - is pooling appropriate, as the means may vary
considerably according to ethnicity?*

We thought about this, and now agree that it is better to use mean
number of children of whites and of blacks separately for each ethnic
group:

> 'RLRS is calculated separately by ethnicity. This is not ideal, because it 
> treats the ethnicities as separate breeding populations. But the alternative of
> calculating pooled RLRS would effectively be estimating natural selection in 
> the whole US population by treating whites as representative. We therefore 
> focus on the white US population, with the caveat that results from this 
> "one data point" may not replicate in other ethnicities or countries.'
